The role of advection versus diffusion in intraventricular transport

### 2025-02-03 paper/main.tex
Discussion 1-3, some Appendix updates

\section*{Discussion}
We have presented a computational model of cerebrospinal fluid (CSF) flow in brain ventricles of 2 days post-fertilization (2dpf) zebrafish embryo.
By imposing tangential traction and normal pressure on the CSF flow, the simulated flow fields had similar features to the flow fields observed experimentally, consisting of vortex-structured flow and pulsatile flow. The CSF flow was used to model transport of a solute by an advection-diffusion equation. Qualitatively, simulations compared well to experiments of Dendra2 fluorescent protein photoconversion. Transport in cilia-mutant fish (\emph{schmalhans} mutants) and a CSF model without cilia, was compared to control experiments and the baseline CSF model, indicating a local effect of cilia motion on protein transport. Using the computational model of flow and transport, we further studied region-specific removal of cilia and ventricles geometry modification, and how these changes affected transport.

\subsection*{The roles of diffusion and advection in transport of photoconverted proteins}
We quantified transport in terms of fluorescence intensity change $\Delta F$ (experiments) and mean concentration $\cbar$ (simulations) as functions of time. For both measures, we observed that transport dynamics in the middle ventricle where Dendra2 fluorescent protein was photoconverted contrasted with the dynamics of the more distal regions of the anterior and posterior ventricles. In experiments, regions distal to the photoconversion site showed an approximately linear increase in $\Delta F$, probably because advection is absent in the transport to the distal regions, such that diffusion is the sole mode of transport. In regions proximal to photoconversion, advection and diffusion occur simultaneously due to the presence of cilia, leading to concave profiles for $\Delta F$. The convex diffusion-only behavior was observed in the mean concentration $\cbar_6$ profile of the posterior ventricle in the simulations, but $\cbar_5$ was concave for the anterior ventricle, opposed to the experimental results \lyng{HH: \emph{Is this the correct interpretation of convex vs. concave curves?} Include some referencing to textbook solutions of the diffusion and advection equations to explain this convex vs. concave? E.g. semi-infinite slab or boundary layer solutions to inflow/outflow problem.}. Based on the geometry modification study, the reason for this difference in behavior might be that the duct between the anterior and middle ventricle is inflated in our computational mesh, as a result of the dye-injection performed before imaging~\lyng{Cite this?}. This inflation allows for effective cilia-mediated advection from the middle ventricle to the anterior ventricle.

In general, the mean concentration dynamics of the simulations was more rapid than the fluorescence intensity dynamics of the photoconversion experiments. This is strongly related to our definition of the photoconversion site (region of interest 1) in the simulations, and whether this corresponds well to the region used in experiments, because the volume of the photoconversion region used in the simulations determines the total amount of solute, and thus impacts the rate of transport. The alignment of the computational and experimental region of interest definition is hard to quantify. Additionally, a source of discrepancy in the results is the fact that the brain ventricles might be permeable, such that some Dendra2 diffuses out of the brain ventricles in the experiments. An assay has shown that the neuroepithelium lining the walls of brain ventricles in embryonic zebrafish 24 hours post-fertilization is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}. Dendra2 are almost 26 kDa in size~\cite{Gurskaya2006EngineeringLight}, and are thus under this reported limit. If such permeability also is the case for 2 days post-fertilization zebrafish embryo ventricles, this would lead to Dendra2 leaking out of the brain ventricles in the experiments, making the dynamics appear slower.

Transport of Dendra2 fluorescent protein appears to be a balance of advection and diffusion. Advection is the strongest transport mechanism on a global scale (global Péclet number 7.4), but for shorter distances diffusion is balancing out advection. Simulating transport of larger molecules (thus with smaller diffusion coefficients) proved advection to be increasingly important as molecule size increased. For the smallest diffusion coefficient used, $D_1$, which represented 200 nm exosomes, there was essentially zero solute in the posterior ventricle at the final simulation time (around 850 s). This is consistent with the statement above that transport to the posterior ventricle from the middle ventricle is driven by diffusion, based on estimated transport times of diffusion and advection: For $D_1$, an estimate of the diffusion timescale $t_d=L^2/D_1$ over a distance $L=200$ $\mu$m from the middle to the posterior ventricle is approximately 24500 s. Advective transport by the mean velocity $\overline{U}=1.41 \ \mu$m/s over the same distance would happen on a timescale of the order of $t_a=L/\overline{U}=141$ s. Comparing this with the minuscule values of $\cbar_6$ after simulating transport with $D_1$ for 850 s, one can conclude that if advection played a role in transporting solute from the photoconversion site in the middle ventricle to the posterior ventricle, values of $\cbar_6$ would be higher than those observed.

The changes in the times to threshold when changing the diffusion coefficient are greatest for regions of interest (ROIs) 4--6, further indicating that, relative to advection, diffusion has a greater impact in these regions than compared to regions 1--3. In region of interest 2 (anterior, dorsal region of the middle ventricle), the time to threshold is lower for $D_2$ than $D_3$. This is probably a result of less solute diffusing away from ROI 1 for $D_2$ than for $D_3$, meaning that more solute can be transported by advection along the ciliated walls of the ventricles towards ROI 2. We also note that, for regions of interest 3 and 4, the order of the $\Delta F$ and $\cbar$ profiles are opposite. This could be related to the difference in the exact location of ROIs 3 and 4 in simulations and experiments, or due to cilia-advection being more prominent in the simulations (cf. the discussion on differences in transport to the anterior ventricle).

\subsection*{Simulation studies: modifying cilia and geometry}
We studied the impacts of removing cilia in specific regions of our model~(\cref{fig:figure5_compare_cilia_modifications}). Removing the dorsal cilia has a strong impact on the resulting CSF flow field and the mean solute concentration. Removing the dorsal cilia decreased the maximum velocity in the ventricles roughly by a factor of three from 27.7 $\mu$m/s to 8.0 $\mu$m/s. The impactful changes to the flow field is because the total force applied on the dorsal cilia boundary is great in magnitude in our model, and the dorsal cilia are thus important to establishing the CSF flow structures. With weaker flow in the dorsal region of the middle ventricle, there is less advective transport. \lyng{HH: Quantify the difference in force applied in the dorsal region: $\int_{c, dorsal}\btau\cdot\mathbf{t}\,\mathrm ds$.}

Absence of ventral cilia in the middle ventricle did not greatly impact flow dynamics and transport, but lead to slightly less transport towards the posterior ventricle, and slightly more transport towards the anterior ventricle. This is because the ventral cilia help induce flow towards the posterior ventricles. Removing the cilia in the anterior ventricle leaves the transport results virtually unaltered, although the CSF flow vortex structure in the anterior ventricle vanishes. This would then probably have a greater impact for smaller diffusion coefficients, as it is evident in~\cref{fig:fig2} that transport with $D_1$ is mainly advective.

@@

Modifying the ventricular geometry had varying impact on transport~(\cref{fig:figure6_sim_results_compare_geometries}). Both when (separately) shrinking the connection between the fore- and mid-brain and the mid- and hind-brain, we observe a slight increase in the mean concentration $\cbar$ of regions of interests 2--4. In regions 5 and 6, we observe a decrease in $\cbar$ in the region that we shrink the connection from the middle ventricle to. Meanwhile, the other region has an increase in $\cbar$, a result of mass conservation. When shrinking the middle ventricle, the mean concentration decreases in regions 2--4. The effect on $\cbar$ in regions 5 and 6 is similar to when the mid-hind connection is shrunk: less solute is transported towards the hind-brain and more solute is transported towards the fore-brain.

Combining all three geometry modifications, we see an increase in $\cbar$ in regions 2--4. Now, there is a decrease in both region 5 and 6, which is more similar to the experimental $\Delta F$ curves (\cref{fig:figure3_compare_exp_and_sim_control}, panel \textbf{a}) in these regions compared to the simulated $\cbar$ on the standard geometry (\cref{fig:figure3_compare_exp_and_sim_control}, panel \textbf{b}). This is in line with the expectation that shrinking the geometry is closer to replicating the geometries of the fish used in experiments, because the ventricular geometry is inflated by the injection procedure performed during image segmentation.

\subsection*{Computational model assumptions and boundary conditions}
To model flow of cerebrospinal fluid, we employed slip boundary conditions at the ventricular walls, contrasting the classical choice of a noslip boundary condition. The use of slip boundary conditions is motivated by experimental observations of the flow field in zebrafish embryo brain ventricles~\cite{Olstad2019CiliaryDevelopment}. It has been observed that the fluid velocity increases towards the wall~(this is also evident in~\cref{fig:figure1}, panel~\textbf{c}). Using slip boundary conditions allows for this behavior. A couple of previous continuum model simulation studies of cilia-mediated flow~\cite{Siyahhan2014FlowVentricles, Thouvenin2020OriginCanal} have modeled cilia as body forces. We chose to model the cilia as a force exerted tangentially on the boundary. First, we deem this a good representation of the real nature of cilia, since they are in reality microscale structures confined to the ventricular walls that generate flow through collective motion. Second, it avoids the need of excessive mesh refinement near the ciliated boundaries in the case that a body force were to be applied in a $\sim$ 5 $\mu$m~\cite{Salman2022ComputationalEmbryo} deep cilia layer.

@@

%%--------------APPENDIX--------------%%
\appendix
\section{Supplementary Theory}
\subsection{Weak Formulation of the Stokes Equations}\label{subsec:appendixA1}
We consider the time-dependent Stokes equations
\begin{subequations}
\begin{alignat}{2}
   \rho\pdifft{\uu} - \nabla \cdot \bsig(\uu, p) &= \mathbf{0} \quad &&\mathrm{in} \ \Omega, \label{eq:BDM_stokes_mom}\\
  \nabla \cdot \uu &= 0 \quad &&\mathrm{in} \ \Omega, \label{eq:BDM_stokes_div} 
\end{alignat}
\label{eq:appendix_stokes_eqs}% 
\end{subequations}%
in the domain $\Omega$ with boundary $\Gamma$, where $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ is the viscous stress tensor, in which $\beps(\uu) = \frac{1}{2}(\nabla\uu + (\nabla\uu)^T)$ is the symmetric gradient. Furthermore, $\uu\in\VV$ is the fluid velocity and $p\in Q$ is the fluid pressure, and $\rho$ and $\mu$ are the fluid density and dynamic viscosity, respectively. Unless otherwise stated, we let $\bsig = \bsig(\uu, p)$.

@@

\iffalse
\textbf{Model Version IIb.} A second approach to modeling the pulsatile flow generated by the cardiac cycle was applying a rigid motion to the geometry, and then solving for the velocity defined in the reference frame attached to the geometry. We consider an approach similar to San Mart{\'\i}n \emph{et al.}~\cite{san2009convergence}, based on the framework commonly referred to as an Arbitrary Lagrangian Eulerian formulation of the governing equations. Consider the Stokes equations 
    \begin{alignat}{2}
        \rho\pdifft{\uu} - \nabla\cdot\bm{\sigma}(\uu, p) &= \mathbf{0}, \\
        \nabla\cdot\uu &= 0,
    \end{alignat}
    \label{eq:stokes_eqs_moving_ref}%
in a time-dependent, moving domain $\Omega_t$, and define the coordinates $\hat{\xx} = (\hat{x}, \hat{y}, \hat{z})$ as the reference frame attached to $\Omega_t$. In $\Omega_t$, we consider the variables $\psi(\hat{\xx}, t)$, so that we have $\uu(\hat{\xx}, t)$ and $p(\hat{\xx}, t)$ in~\eqref{eq:stokes_eqs_moving_ref}. As before, we denote $\Omega$ the domain in the absolute, fixed reference frame with coordinates $\xx = (x, y, z)$ and variables $\psi(\xx, t)$. To determine the velocity inside the brain ventricles without moving the computational mesh, we can solve the Stokes equations for the velocity $\ww$ defined in the moving reference frame. This is done by first solving for $\uu$ in the fixed reference frame, whilst accounting for the mesh motion $\uuref$. Then, with $\uu$ and $\ww$ defined in separate coordinate systems that move with a velocity $\uuref$ relative to each other, the velocities are related by $\uu = \ww + \uuref$~\cite{meriam2012engineering}. After solving for $\uu$, we thus calculate $\ww = \uu - \uuref$.  To rewrite the variables in~\eqref{eq:stokes_eqs_moving_ref} into variables of $\xx$ instead of $\hat{\xx}$, nothing has to be done with the divergence operators since we consider a rigid motion, rendering spatial differentiation unchanged. The temporal derivative is not the same in the two coordinate systems, however. Differentiation with respect to time and application of the chain rule yields
    \begin{aligned}
        \pdifft{\psi}(\hat{\xx}, t) &= \pdifft{\psi}(\xx, t) + \frac{\partial \psi}{\partial\hat{\xx}}\pdifft{\hat{\xx}} \\
        &= \pdifft{\psi}(\xx, t) + \frac{\partial \psi}{\partial\xx}\frac{\partial\xx}{\partial\hat{\xx}}\pdifft{\hat{\xx}} \\
        &= \pdifft{\psi}(\xx, t) + \frac{\partial \psi}{\partial\xx}\pdifft{\hat{\xx}}.
    \end{aligned}
The last equality follows from the fact that the motion of the moving reference frame is rigid, such that the differences in $\hat{\xx}$ and $\xx$ are only due to translation, and not dilatation, of the coordinate systems. We can recognize $\partial\hat{\xx}/\partial t = \uu_{\mathrm{ref}}$ as the velocity of the domain $\Omega_t$. Since $\partial\psi/\partial\xx = \nabla\psi$, we get
    \pdifft{\psi}(\hat{\xx}, t) = \pdifft{\psi}(\xx, t) + \uu_{\mathrm{ref}}\cdot\nabla\psi,
    \label{eq:temporal_derivative_relation}
where the second term leads to the interpretation that the variable $\psi(\xx, t)$ in the fixed reference frame is advected with the velocity $\uu_{\mathrm{ref}}$ of the moving reference frame. Using the relation~\eqref{eq:temporal_derivative_relation}, we can rewrite~\eqref{eq:stokes_eqs_moving_ref} in the fixed reference frame:
    \begin{alignat}{2}
        \rho\bigg(\pdifft{\uu} + (\uu_{\mathrm{ref}}\cdot\nabla)\uu\bigg) &= \nabla\cdot\bm{\sigma}(\uu, p), \label{eq:mom_eq_fixed}\\
        \nabla\cdot\uu &= 0,
    \end{alignat}
    \label{eq:stokes_eqs_fixed}%
where the variables in~\eqref{eq:stokes_eqs_fixed} are variables of $\xx$ and $t$. The boundary conditions also change slightly for this formulation. Since the impermeability condition $\ww\cdot\nn=0$ is now supposed to hold in the moving reference frame, we get the boundary condition $\uu\cdot\nn = \uuref\cdot\nn$. To simulate the oscillatory motion arising from the cardiac cycle, we applied the motion $\uuref = (A_{\mathrm{ref}}\sin{\omega t}, 0, 0)$, where the amplitude $A_{\mathrm{ref}}$ was calibrated with experimental data and $\omega = 2\pi f$ was determined by the cardiac frequency $f$~\cite{Olstad2019CiliaryDevelopment}. Since the domain movement is rigid, we have $\beps(\uuref) = \mathbf{0}$, such that the stress boundary condition $P_{\nn}(\bsig(\ww, p)) = \mathbf{0}$ directly translates to $P_{\nn}(\bsig(\uu, p)) = \mathbf{0}$. Solving~\eqref{eq:stokes_eqs_fixed} with the boundary conditions $\uu\cdot\nn = \uuref\cdot\nn$ and $P_{\nn}(\bsig(\uu, p)) = \mathbf{0}$, we obtain the velocity in the moving reference frame by calculating $\ww = \uu - \uuref$.
\fi

@@

\subsection{Discretization of the Stokes Equations}
We will outline the discretization of the weak formulation~\cref{eq:abstract_weak_form_modelC,eq:stokes_bilinear_form,eq:stokes_skew_form,eq:stokes_linear_form} of the Stokes equations for model version 0. A similar approach can be used to discretize the weak formulation for the other model versions. First, we consider the spatial discretization of the weak formulation. We rewrite the weak form in terms of the discretized variables, denoted with the subscript $h$:
\begin{subequations}
    \begin{alignat}{2}
        \mathcal{A}(\uu_h, \vv_h) + \mathcal{B}(\vv_h, p_h) &= \mathcal{L}(\vv_h), &&\quad\forall\vv_h\in\VV_h, \\
        \mathcal{B}(\uu_h, q_h) &= 0, &&\quad\forall q_h\in Q_h.
    \end{alignat}%
    \label{eq:no_stabilization_abstract_discrete_weak_form_modelC}%
\end{subequations}%
Our choice for the discrete spaces $\VV_h$ and $Q_h$ are finite elements of the Brezzi-Douglas-Marini family~\cite{Brezzi1985TwoProblems} and discontinuous Galerkin elements, respectively. Application of this scheme requires a stabilization of the bilinear form $\mathcal{A}(\uu_h, \vv_h)$. We apply the stabilization outlined in Hong~\emph{et al.}~\cite{Hong2016AEquations}. First, we define the average and jump operators, respectively, as
\begin{equation*}
    \avg{\uu} = \frac{\uu^+\cdot\nn^+ - \uu^-\cdot\nn^-}{2} \qquad\mathrm{and}\qquad \jump{\uu} = \uu^+ - \uu^-,
\end{equation*}
where $\nn$ is the normal vector on a facet separating two cells, where the plus and minus signs denote variables in each of the two cells. The normal vector is chosen such that it points towards the exterior of its corresponding cell. On the boundary, the operators are defined as $\avg{\uu} = \uu\cdot\nn$ and $\jump{\uu} = \uu$,
with $\nn$ pointing to the exterior of the domain. We now define and add the bilinear form
\begin{align*}
    \mathcal{A}_{\mathrm{stab}}(\uu_h, \vv_h) &= \sum_{k\in K}\intF{2\mu\frac{\gamma}{\avg{h}}\jump{P_{\nn}(\uu_h)}\cdot \jump{P_{\nn}(\vv_h)}}\\
    &-\sum_{k\in K}\intF{2\mu\avg{\beps(\uu_h)}\cdot\jump{\vv_h}}-\sum_{k\in K}\intF{2\mu\avg{\beps(\vv_h)}\cdot\jump{\uu_h}}
\end{align*}
to the discrete weak  formulation~\eqref{eq:no_stabilization_abstract_discrete_weak_form_modelC}, where $\mathcal{F} = \{\mathcal{F}_k : k\in K\}$ is the set of all interior facets $\mathcal{F}_k$ of the computational mesh. The parameter $\gamma>0$ is introduced to penalize jumps in the tangential components of the velocity across facets, wheras $h$ is the cell diameter of a mesh cell, defined as the maximal distance between two points in a cell. The semi-discrete weak formulation of the Stokes equations reads: find ($\uu_h, p_h$)$\in$($\VV_h, Q_h$) such that
\begin{subequations}
    \begin{alignat}{2}
        \mathcal{A}(\uu_h, \vv_h) + \mathcal{A}_{\mathrm{stab}}(\uu_h, \vv_h) + \mathcal{B}(\vv_h, p_h) &= \mathcal{L}(\vv_h), &&\quad\forall\vv_h\in\VV_h, \\
        \mathcal{B}(\uu_h, q_h) &= 0, &&\quad\forall q_h\in Q_h.
    \end{alignat}%
    \label{eq:abstract_semi_discrete_weak_form_modelC}%
\end{subequations}%
We solved~\eqref{eq:abstract_semi_discrete_weak_form_modelC} numerically with the penalty parameter $\gamma=10$.

@@

\subsection{Weak Formulation of the Advection-Diffusion Equation}
Recall the strong form of the advection-diffusion equation:
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad\mathrm{in} \ \Omega,
    \label{eq:adv_diff_strong_DG}
\end{equation}
where $\Omega$ is the domain with boundary $\Gamma$, and the total solute flux
\begin{equation*}
    \JJ(c) = c\uu - D\nabla c.
\end{equation*}
We consider $\JJ = \JJ(c)$, unless otherwise indicated. The concentration $c\in W$ and the cerebrospinal fluid velocity $\uu\in\VV$. The weak form of~\eqref{eq:adv_diff_strong_DG} is derived by multiplying the equation with a test function $\phi \in W$ and integrating over the domain: 
\begin{equation*}
    \intO{\pdifft{c}\phi} + \intO{(\nabla\cdot\JJ)\phi} = 0.
    %\label{eq:adv_diff_weak_DG_1}
\end{equation*}
Integrating the second term by parts and applying Green's first identity yields
\begin{equation}
    \intO{\pdifft{c}\phi} + \intG{(\JJ\cdot\nn)\phi} - \intO{\JJ\cdot\nabla\phi} = 0.\label{eq:adv_diff_weak_formA1}
\end{equation}
The boundary integral of the total flux is handled differently for the three different model versions, depending on the boundary conditions considered.

>>>>

\section*{Discussion}
We have presented a computational model of CSF flow and solute transport within the brain ventricles.
In this study, we used the larval zebrafish brain ventricles since its geometry,
flow components and kinetics are well described, thereby allowing us to validate
our simulation experimentally. Nevertheless, our simulations can be implemented in
other systems in the future when their flow parameters are comprehensively resolved.  
By imposing tangential traction and normal pressure on the CSF flow, we achieved to reproduce
flow fields with similar features as observed experimentally, consisting of vortex-structured flow,
mediated by cilia, and pulsatile flow, mediated by the cardiac cycle.

Using the simulated CSF flow, we modeled solute transport by an advection-diffusion equation.
These results compared well with experimental data performed \emph{in vivo} in zebrafish.
Next, by using both a model and zebrafish mutants with paralyzed cilia,
we identified that even though advection by cilia is dominant,
diffusion plays a significant role  in protein transport. Finally, we report how local
flow patterns and ventricular geometries affect solute distribution.
Altogether our results provide not only new findings on the role of
diffusion and advection but also offer a new framework to study CSF transport in complex geometries.

\subsection*{The roles of advection and diffusion in intraventricular solute transport}
To determine the roles of advection and diffusion in solute transport, we quantified transport
based on changes in fluorescence intensity $\Delta F$ (experiments) and
mean concentration $\cbar$ (simulations) as functions of time.
Even though we identified very similar dynamics in simulations and
\emph{in vivo} experiments, the dynamics of the simulations were generally
more rapid and resulted in higher total amounts.
Differences between experiments and simulations may relate to
ventricular size, as we showed that modifying the duct volumes impacts
solute distribution and the geometry used for the model was larger than in experiments.
Alternatively, it may relate to the way we defined the photoconversion site (region of interest 1)
in the simulations, since the volume of the photoconversion region determines
the total amount of solute, and thereby the rate of transport. Another source
of discrepancy may relate to the permeability of the ventricular wall,
which we have ignored in our simulation. Reports have shown that the
neuroepithelium lining the walls of brain ventricles in 24h old zebrafish
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
Since Dendra2 is around 26 kDa in size~\cite{Gurskaya2006EngineeringLight},
and thus under this reported limit, one could assume that if Dendra2
leaks out of the brain ventricles, the dynamics become slower.

In general, the mean concentration dynamics of the simulations was 
more rapid than the fluorescence intensity dynamics of the photoconversion experiments.
This is strongly related to our definition of the photoconversion site
(ROI 1) in the simulations, and whether this corresponds well to the
region used in experiments, because the volume of the photoconversion
region used in the simulations determines the total amount of solute,
and thus impacts the rate of transport. The alignment of the computational
and experimental region of interest definition is hard to quantify.
Additionally, a source of discrepancy in the results is the fact that the
brain ventricles might be permeable, such that some Dendra2 diffuses out
of the brain ventricles in the experiments. An assay has shown that the neuroepithelium
lining the walls of brain ventricles in embryonic zebrafish 24 hours post-fertilization
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
Dendra2 are almost 26 kDa in size~\cite{Gurskaya2006EngineeringLight},
and are thus under this reported limit. If such permeability also is the case for 2 days
post-fertilization zebrafish embryo ventricles, this would lead to Dendra2 leaking
out of the brain ventricles in the experiments, making the dynamics appear slower.

Nevertheless, our results convincingly show that intraventricular transport of proteins
depends on a balance of advection and diffusion. While advection is the strongest transport
mechanism on a global scale for Dendra2 (global Péclet number 7.4), diffusion is balancing out
advection especially for shorter distances and for regions devoid of motile cilia activity.
Indeed, we identified that the diffusion coefficient had a greater impact on transport to
regions located upstream in the CSF flow field (ROI 4\lyng{ actually 3?}) or in different ventricles (ROIs 5--6),
than the regions located downstream of cilia-mediated flow (ROI 2).

In contrast, we identified that transport of larger molecules
(thus with smaller diffusion coefficients),
such as 200 nm exosomes, are almost entirely dependent on advection.
Indeed, there was essentially zero solute in the posterior ventricle at the
final simulation time (around 850 s), where transport is largely driven by diffusion.
For $D_1$, an estimate of the diffusion timescale $t_d=L^2/D_1$ over a distance
$L=200$ $\mu$m from the middle to the posterior ventricle is approximately 24500 s.
Advective transport by the mean velocity $\overline{U}=1.41 \ \mu$m/s over the same distance would happen
on a timescale of the order of $t_a=L/\overline{U}=141$ s. Comparing this with the minuscule values of
$\cbar_6$ after simulating transport with $D_1$ for 850 s, one can conclude that if advection
played a role in transporting solute from the photoconversion site in the middle ventricle
to the posterior ventricle, values of $\cbar_6$ would be higher than those observed.

\subsection*{Impact of cilia properties and ventricular geometry}
To gain further insights into how CSF distribution depends on cilia properties and geometries,
we manipulated certain parameters in our models. First, we removed cilia in specific regions of
our model (\Cref{fig:figure5_compare_cilia_modifications}). We identified that the cilia in the
dorsal middle ventricle contribute the most to solute distribution, which aligns well with the fact
that these cilia generate the strongest flow field \emph{in vivo} and in our model.
In contrast, removing the ventral cilia in the middle ventricle did not greatly impact flow dynamics.
However, it lead to reduced transport towards the posterior ventricle,
and increased transport towards the anterior ventricle, as the ventral cilia help induce
flow towards the posterior ventricles. Removing the cilia in the anterior ventricle did not affect the transport.

Next, we modified the ventricular geometry and identified varying impact on transport.
Notably, shrinking ducts reduced movement between the ventricles, while shrinking the middle ventricle
reduced diffusive transport upstream of the flow field (ROI 4).
We previously identified large inter-individual variation in ventricular size and
CSF velocities between larva at given developmental stages and throughout development~\fixme{Add some citation here? Olstad/D'Gama}.
Our results suggest that these two parameters may well co-vary and explain the variation.

Altogether, our model is flexible and thereby allows \emph{in silico} testing of
further hypotheses on the contribution of ciliated lineages and geometries.
We trust that it will be useful as a basis for future analysis of CSF dynamics in more complex geometries.

@@

\subsection*{Computational model assumptions and boundary conditions}
To model CSF, we employed slip boundary conditions at the ventricular walls,
contrasting the classical choice of a noslip boundary condition.
The use of slip boundary conditions is motivated by experimental observations of the
flow field in zebrafish embryo brain ventricles~\cite{Olstad2019CiliaryDevelopment}.
It has been observed, both previously and in this work, that the fluid velocity increases towards the wall~(\Cref{fig:figure1}c).
Using slip boundary conditions allows for this behavior.
A couple of previous continuum model simulation studies of cilia-mediated flow~\cite{Siyahhan2014FlowVentricles, Thouvenin2020OriginCanal}
have modeled cilia as body forces. We chose to model the cilia as a force exerted tangentially on the boundary.
First, we deem this a good representation of the real nature of cilia,
since they are in reality microscale structures confined to the ventricular walls that
generate flow through collective motion. Second, it avoids the need of excessive mesh
refinement near the ciliated boundaries in the case that a
body force were to be applied in a $\sim$ 5 $\mu$m~\cite{Salman2022ComputationalEmbryo} deep cilia layer.

@@

%%--------------APPENDIX--------------%%
\appendix
\section{Supplementary Theory}
\subsection{Weak formulation of the Stokes equations}\label{subsec:appendixA1}
We consider the time-dependent Stokes equations
\begin{subequations}
\begin{alignat}{2}
   \rho\pdifft{\uu} - \nabla \cdot \bsig(\uu, p) &= \mathbf{0} \quad &&\mathrm{in} \ \Omega, \label{eq:BDM_stokes_mom}\\
  \nabla \cdot \uu &= 0 \quad &&\mathrm{in} \ \Omega, \label{eq:BDM_stokes_div} 
\end{alignat}
\label{eq:appendix_stokes_eqs}% 
\end{subequations}%
in the domain $\Omega$ with boundary $\Gamma$, where $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ is the viscous stress tensor, in which $\beps(\uu) = \frac{1}{2}(\nabla\uu + (\nabla\uu)^T)$ is the symmetric gradient. Furthermore, $\uu\in\VV$ is the fluid velocity and $p\in Q$ is the fluid pressure, and $\rho$ and $\mu$ are the fluid density and dynamic viscosity, respectively. Unless otherwise stated, we let $\bsig = \bsig(\uu, p)$.

@@

@@

\subsection{Discretization of the Stokes equations}
We will outline the discretization of the weak formulation~\cref{eq:abstract_weak_form_modelC,eq:stokes_bilinear_form,eq:stokes_skew_form,eq:stokes_linear_form} of the Stokes equations for model version 0. A similar approach can be used to discretize the weak formulation for the other model versions. First, we consider the spatial discretization of the weak formulation. We rewrite the weak form in terms of the discretized variables, denoted with the subscript $h$:
\begin{subequations}
    \begin{alignat}{2}
        \mathcal{A}(\uu_h, \vv_h) + \mathcal{B}(\vv_h, p_h) &= \mathcal{L}(\vv_h), &&\quad\forall\vv_h\in\VV_h, \\
        \mathcal{B}(\uu_h, q_h) &= 0, &&\quad\forall q_h\in Q_h.
    \end{alignat}%
    \label{eq:no_stabilization_abstract_discrete_weak_form_modelC}%
\end{subequations}%
Our choice for the discrete spaces $\VV_h$ and $Q_h$ are finite elements of the Brezzi-Douglas-Marini family~\cite{Brezzi1985TwoProblems} and discontinuous Galerkin elements, respectively. Application of this scheme requires a stabilization of the bilinear form $\mathcal{A}(\uu_h, \vv_h)$. We apply the stabilization outlined in Hong~\emph{et al.}~\cite{Hong2016AEquations}. First, we define the average and jump operators, respectively, as
\begin{equation*}
    \avg{\uu} = \frac{\uu^+\cdot\nn^+ - \uu^-\cdot\nn^-}{2} \qquad\mathrm{and}\qquad \jump{\uu} = \uu^+ - \uu^-,
\end{equation*}
where $\nn$ is the normal vector on a facet separating two cells, where the plus and minus signs denote variables in each of the two cells. The normal vector is chosen such that it points towards the exterior of its corresponding cell. On the boundary, the operators are defined as $\avg{\uu} = \uu\cdot\nn$ and $\jump{\uu} = \uu$,
with $\nn$ pointing to the exterior of the domain. We now define and add the bilinear form
\begin{align*}
    \mathcal{A}_{\mathrm{stab}}(\uu_h, \vv_h) &= \sum_{k\in K}\intF{2\mu\frac{\gamma}{\avg{h}}\jump{P_{\nn}(\uu_h)}\cdot \jump{P_{\nn}(\vv_h)}}\\
    &-\sum_{k\in K}\intF{2\mu\avg{\beps(\uu_h)}\cdot\jump{\vv_h}}-\sum_{k\in K}\intF{2\mu\avg{\beps(\vv_h)}\cdot\jump{\uu_h}}
\end{align*}
to the discrete weak  formulation~\eqref{eq:no_stabilization_abstract_discrete_weak_form_modelC}, where $\mathcal{F} = \{\mathcal{F}_k : k\in K\}$ is the set of all interior facets $\mathcal{F}_k$ of the computational mesh. The parameter $\gamma>0$ is introduced to penalize jumps in the tangential components of the velocity across facets, wheras $h$ is the cell diameter of a mesh cell, defined as the maximal distance between two points in a cell. The semi-discrete weak formulation of the Stokes equations reads: find ($\uu_h, p_h$)$\in$($\VV_h, Q_h$) such that
\begin{subequations}
    \begin{alignat}{2}
        \mathcal{A}(\uu_h, \vv_h) + \mathcal{A}_{\mathrm{stab}}(\uu_h, \vv_h) + \mathcal{B}(\vv_h, p_h) &= \mathcal{L}(\vv_h), &&\quad\forall\vv_h\in\VV_h, \\
        \mathcal{B}(\uu_h, q_h) &= 0, &&\quad\forall q_h\in Q_h.
    \end{alignat}%
    \label{eq:abstract_semi_discrete_weak_form_modelC}%
\end{subequations}%
We solved~\eqref{eq:abstract_semi_discrete_weak_form_modelC} numerically with the penalty parameter $\gamma=10$.

@@

\subsection{Weak formulation of the advection-diffusion equation}
Recall the strong form of the advection-diffusion equation:
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad\mathrm{in} \ \Omega,
    \label{eq:adv_diff_strong_DG}
\end{equation}
where $\Omega$ is the domain with boundary $\Gamma$, and the total solute flux
\begin{equation*}
    \JJ(c) = c\uu - D\nabla c.
\end{equation*}
We consider $\JJ = \JJ(c)$, unless otherwise indicated. The concentration $c\in W$ and the cerebrospinal fluid velocity $\uu\in\VV$. The weak form of~\eqref{eq:adv_diff_strong_DG} is derived by multiplying the equation with a test function $\phi \in W$ and integrating over the domain: 
\begin{equation*}
    \intO{\pdifft{c}\phi} + \intO{(\nabla\cdot\JJ)\phi} = 0.
    %\label{eq:adv_diff_weak_DG_1}
\end{equation*}
Integrating the second term by parts and applying Green's first identity yields
\begin{equation}
    \intO{\pdifft{c}\phi} + \intG{(\JJ\cdot\nn)\phi} - \intO{\JJ\cdot\nabla\phi} = 0.\label{eq:adv_diff_weak_formA1}
\end{equation}
The boundary integral of the total flux is handled differently for the three different model versions, depending on the boundary conditions considered.
